# Reproducible Research: Peer Assessment 1
======================================================

## step 1. Load the data 

```{r echo =TRUE}
activity.data <- read.csv("activity.csv")
apply(activity.data, 2, class)
summary(activity.data)
```

Missing values: How many records with missing data?
total number of records, missing values for each variable
```{r}
nrow(activity.data); sum(is.na(activity.data$steps)); sum(is.na(activity.data$date)); sum(is.na(activity.data$interval))
```

## step 2 preprocess the data
convert steps to numeric might be useful later.
convert date to type Date
```{r}
activity.data$steps <- as.numeric(activity.data$steps)
activity.data$date <- as.Date(activity.data$date)
activity.data$interval <- as.integer(activity.data$interval)
class(activity.data$steps)
class(activity.data$date)
str(activity.data)
```

# Key results

## What is mean total number of steps taken per day?

the mean number of steps taken per day regardless of missing values:
```{r}
sum(activity.data$steps, na.rm=TRUE)/61
```
Histogram of the total number of steps taken each day
```{r}
sum.steps.day <- aggregate(activity.data$steps, list(activity.data$date), sum)
library(lattice)
histogram(~x, data = sum.steps.day, xlab = "steps",ylab = "freqeuncy - days", main = "Histogram of total number of steps taken each day" )
```

mean steps taken per day
```{r}
steps.day <- aggregate(activity.data$steps, list(activity.data$date), sum)
head(steps.day)

mean(steps.day$x, na.rm=TRUE)
```
median steps taken per day
```{r}
median(steps.day$x, na.rm=TRUE)
```


## What is the average daily activity pattern?
```{r}
complete.data <- activity.data[complete.cases(activity.data), ]
sum(is.na(complete.data$steps))  ## check that NAs hv been removed
nrow(complete.data)
steps.interval <- aggregate(complete.data$steps, list(complete.data$interval), mean)
colnames(steps.interval) <- c("interval","mean.steps")
nrow(steps.interval)
str(steps.interval)
xyplot(mean.steps~interval, steps.interval, type ="l")
```

the 5 minute interval that accounts for the max no of steps on average across all days
```{r}
tail(steps.interval[order(steps.interval$mean.steps), ], 1)
```

Answer: The 0835 to 0840 5 minute interval sees the most steps This is also shown in the time series plot.

## Imputing missing values
My strategy is to assume that activity level in a given 5 minute interval is pretty much the same each day.
very likely there is some variation between weekday vs weekend and even across the day of the week
But to kep it simple I have assumed that it would be reasonable to replace a missing value with the mean for 
that timeslot across all the days for which there is data. i.e I have not made any adjustments for w/day / w/end.
The following function does the imnputatation

```{r}
replaceNA <- function(activity.data) {
   for (i in 1 : nrow(activity.data)) {
      if (is.na(activity.data[i, 1])) {
        for (j in 1 : nrow(steps.interval)) {
          if (activity.data[i, 3] == steps.interval[j, 1]) {
            #print(i); print(j)
            #print(steps.interval[j, 2])
            activity.data[i, 1] <- steps.interval[j, 2]
            #print(activity.data[i, 1])
          } # end of inner if
        } # end of inner for loop
      } #end if
   }#end of 1st for loop
return(activity.data)
} # endfunction

```

Run the replaceNA function
```{r}
imputed.data <- replaceNA(activity.data)

```

Check that it has worked There should be no NAs in imputed.data

```{r}
head(imputed.data)
tail(imputed.data)
sum(is.na(imputed.data$steps))
```

It works!!

## Are there differences in activity patterns between weekdays and weekends?

add a variable to the dataframe imputed.data to show whether its a weekend or a weekday

```{r}
str(imputed.data)
```

```{r}
fun.days <- c("Saturday", "Sunday")
imputed.data$weekend <- ifelse(weekdays(imputed.data$date) %in% fun.days, c("Weekend"),
                               c("Weekday") ) 
str(imputed.data)
imputed.data$weekend <- as.factor(imputed.data$weekend)
table(imputed.data$weekend)
library(lattice)
attach(imputed.data)
xyplot(steps~interval|weekend, type = "l", 
       main = "Activity by time slots",
       layout = c(1,2))
```

Panel plot of steps against time slots to show weekday / weekend differences.
My conclusion: Weekday activity starts much earlier in the day. On weekends starts laterand appears to peak late afternoon / early evening.



