# Reproducible Research: Peer Assessment 1
======================================================

## step 1. Load the data 

```{r echo =TRUE}
activity.data <- read.csv("activity.csv")
apply(activity.data, 2, class)
summary(activity.data)
```

Missing values: How many records with missing data?
total number of records, missing values for each variable
```{r}
nrow(activity.data); sum(is.na(activity.data$steps)); sum(is.na(activity.data$date)); sum(is.na(activity.data$interval))
```

## step 2 preprocess the data
convert steps to numeric might be useful later.
convert date to type Date
```{r}
activity.data$steps <- as.numeric(activity.data$steps)
activity.data$date <- as.Date(activity.data$date)
class(activity.data$steps)
class(activity.data$date)
```

# Key results

## What is mean total number of steps taken per day?

the mean number of steps taken per day regardless of missing values:
```{r}
sum(activity.data$steps, na.rm=TRUE)/61
```
Histogram of the total number of steps taken each day
```{r}
sum.steps.day <- aggregate(activity.data$steps, list(activity.data$date), sum)
library(lattice)
histogram(~x, data = sum.steps.day, xlab = "steps",ylab = "freqeuncy - days", main = "Histogram of total number of steps taken each day" )
```

mean steps taken per day
```{r}
steps.day <- aggregate(activity.data$steps, list(activity.data$date), sum)
mean(steps.day$x, na.rm=TRUE)
```
median steps taken per day
```{r}
median(steps.day$x, na.rm=TRUE)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
